---
title: "UF - Morrison Lab Preliminary Piney Point Results"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

Preliminary analysis of results from NSF RAPID project on tracking the effects of Piney Point discharge on Tampa Bay. Sites throughout Tampa Bay and St. Joseph Sound were sampled at minimum every two weeks. 

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(cowplot)
library(knitr)

knitr::opts_chunk$set(echo = TRUE)


```

# Bulk Particulate Organic Matter
```{r, include=FALSE, echo=FALSE}
raw_POM<-readxl::read_excel("../Data/UF_SIL/Full_PP_GFF_data.xlsx", sheet = 1)


noNA_date<-raw_POM[!is.na(raw_POM$Date), ]  
noNA_date$`wt %N`<-as.numeric(noNA_date$`wt %N`)
noNA_date$`wt %C`<-as.numeric(noNA_date$`wt %C`)

noNA_date<-mutate(noNA_date, CN=`wt %C`/`wt %N`)
noNA_onlysites<-filter(noNA_date, Location!=("Blank")&Location!=("Effluent"))
effluent<-filter(noNA_date, grepl("filter_pineypoint3", GFF_ID))


```


### d13C for all locations and sampling dates
```{r, echo=FALSE, warning=FALSE}
eff_d13C<-ggplot(effluent, aes(Location, `d13C (permil, vs VPDB)`))+
  geom_point()+
  scale_y_continuous(limits = c(-31, -15))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

POM_d13C<-ggplot(noNA_onlysites, aes(Date, `d13C (permil, vs VPDB)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

plot_grid(POM_d13C, eff_d13C, rel_widths = c(1, 0.25))
```
<!---
Figure 1. Changes in d13C values for particulate organic matter (POM) over time. There is a clear enrichment of d13C values over the course of the study period. Protein/carbohydrates higher/more enriched d13C value, lipids/lignin lower/depleted(more negative) value. OM decomposition selectively removes labile 13C depleting d13C. Photosynthetic carbon (d13C -27permil), atmospheric C (d13C -7permil). Photosynthetic biomass enriched, since dissolved CO2 or HCO3- is used for C fixation(d13C = 0permil). C4 plants more enriched (d13C values between 8-18permil), Algae pump either CO2 or HCO3 as a C source for photosynthesis (d13C = 0), while land plants use CO2 from atmosphere (d13C= -7permil), so algae have more enriched signature. (From Bianchi and Canuel, 2011). Effluent POM could be a combination of C4 plants (-16 to -9 permil; Sharp 2017), higher plants (-11.54 to -13.21permil, Homosassa to Wacassasa; Barry et al., 2018), algae -23.4permil; sediment (-16.14 to -19.86), seagrass leaf d13C was between -12 and -5permil. 
-->

```{r, eval=FALSE, echo=FALSE}
#double check GFF ids
noPP<-filter(noNA_onlysites, !grepl("PP", GFF_ID))

noPP_d13C<-ggplot(noPP, aes(Date, `d13C (permil, vs VPDB)`, color=GFF_ID))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

noPP_d15N<-ggplot(noPP, aes(Date, `d15N (permil, vs AIR)`, color=GFF_ID))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")
```


### d15N for all locations and sampling dates
```{r, echo=FALSE, warning=FALSE}

eff_d15N<-ggplot(effluent, aes(Location, `d15N (permil, vs AIR)`))+
  geom_point()+
  scale_y_continuous(limits = c(-18, 4))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

POM_d15N<-ggplot(noNA_onlysites, aes(Date, `d15N (permil, vs AIR)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

plot_grid(POM_d15N, eff_d15N, rel_widths = c(1, 0.25))
```
<!---Figure 2. POM d15N values. Inorganic N species can have wide range of d15N values. d15N enriched 2 to 3 permil with each trophic level (Bianchi and Canuel, 2011). From Sharp (2017) average value of N fixation is less than 0permil. From Sharp 2017/Velinsky et al (1991), waters with NH4+ of 40uM, fractionation between POM and NH4 modeled to be -20 to -30permil. with lower concentrations, 9uM fractionations were -15 to -15permil. **Effluent had a NH4-N concentration of 61.6 uM.** Ammonium is lighter than nitrate by several per mil. Peterson and Fry's average estimate for precip are -18 to 8 for ammonium and -15 to 3 permil for NO3-. d15N values from most ocean materials are positive. 
-->

### C:N for all locations and sampling dates
```{r, echo=FALSE, warning=FALSE}

eff_CN<-ggplot(effluent, aes(Location, CN))+
  geom_point()+
  scale_y_continuous(limits = c(3, 13))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

POM_CN<-ggplot(noNA_onlysites, aes(Date, CN))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

plot_grid(POM_CN, eff_CN, rel_widths = c(1, 0.25))
```

### %TN for all locations and sampling dates
```{r, echo=FALSE, warning=FALSE}

eff_TN<-ggplot(effluent, aes(Location, `wt %N`))+
  geom_point()+
  scale_y_continuous(limits = c(0, 3))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

POM_TN<-ggplot(noNA_onlysites, aes(Date, `wt %N`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

plot_grid(POM_TN, eff_TN, rel_widths = c(1, 0.25))
```

### %TC for all locations and sampling dates
```{r, echo=FALSE, warning=FALSE}

eff_TC<-ggplot(effluent, aes(Location, `wt %C`))+
  geom_point()+
  scale_y_continuous(limits = c(1, 12))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

POM_TC<-ggplot(noNA_onlysites, aes(Date, `wt %C`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")

plot_grid(POM_TC, eff_TC, rel_widths = c(1, 0.25))
```




```{r, include=FALSE, echo=FALSE}
###extra code for shading dates
 # geom_rect(data = data.frame(xmin = decimal_date(as.Date(c("1924-01-01"))),
 #                              xmax = decimal_date(as.Date(c("1928-12-31"))),
 #                              ymin = -Inf,
 #                              ymax = Inf),
 #            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
 #            fill = "grey", alpha = 0.5)

```

## Nutrient data
```{r, echo=FALSE, warning=FALSE, include=FALSE}

nutrients_raw<-readxl::read_excel("../Data/PineyPoint_all_samples_w_data.xlsx")

nutrients_raw$`OrthoP (ug/L)`<-as.numeric(nutrients_raw$`OrthoP (ug/L)`)

nutrients_onlysites<-filter(nutrients_raw, Location!=("Blank")&Location!=("Effluent"))
nutrients_effluent<-filter(nutrients_raw, Location==("Effluent"))


```

### Orthophosphate
```{r, echo=FALSE, warning=FALSE}

eff_orthoP<-ggplot(nutrients_effluent, aes(Location, `OrthoP (ug/L)`))+
  geom_point()+
  scale_y_continuous(limits = c(0, 170000))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

orthoP<-ggplot(nutrients_onlysites, aes(`Date of Sample collection`, `OrthoP (ug/L)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
  panel_border(color="black")

plot_grid(orthoP, eff_orthoP, rel_widths = c(1, 0.25))
```

### Total phosphorus
```{r, echo=FALSE, warning=FALSE}

eff_TP<-ggplot(nutrients_effluent, aes(Location, `TotalP (ug/L)`))+
  geom_point()+
  scale_y_continuous(limits = c(0, 170000))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

TP<-ggplot(nutrients_onlysites, aes(`Date of Sample collection`, `TotalP (ug/L)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
  panel_border(color="black")

plot_grid(TP, eff_TP, rel_widths = c(1, 0.25))
```

### NH4-N
```{r, echo=FALSE, warning=FALSE}

eff_NH4<-ggplot(nutrients_effluent, aes(Location, `NH4N (mg/L)`))+
  geom_point()+
  scale_y_continuous(limits = c(0, 2))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

nutrients_onlysites_greater0_NH4<-filter(nutrients_onlysites, `NH4N (mg/L)`>0)

NH4<-ggplot(nutrients_onlysites_greater0_NH4, aes(`Date of Sample collection`, `NH4N (mg/L)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
  panel_border(color="black")

plot_grid(NH4, eff_NH4, rel_widths = c(1, 0.25))
```

### NOx
```{r, echo=FALSE, warning=FALSE}

eff_NOx<-ggplot(nutrients_effluent, aes(Location, `NOx (mg/L)`))+
  geom_point()+
  scale_y_continuous(limits = c(0, 1))+
  theme_cowplot(font_size = 8)+
   panel_border(color="black")+
  xlab("")

#remove values <0, because below LOD
NOx_data<-filter(nutrients_onlysites, `NOx (mg/L)`>0)

NOx<-ggplot(NOx_data, aes(`Date of Sample collection`, `NOx (mg/L)`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
  panel_border(color="black")

plot_grid(NOx, eff_NOx, rel_widths = c(1, 0.25))
```

```{r, echo=FALSE, warning=FALSE}

chl<-ggplot(nutrients_onlysites, aes(`Date of Sample collection`, `Total Chl`))+
  geom_point()+
  facet_wrap(.~Location)+
  theme_cowplot(font_size = 8)+
  panel_border(color="black")+
  ylab("Total chlorophyll (ug/L)")

chl
```
